\textsc{Rils}-\textsc{Rols} : Robust Symbolic Regression via Iterated Local Search and Ordinary Least Squares

### 2022-11-08 paper_resources/doc/paper.tex
Chapetr 3: polished

\fxnote{Dati neku smislenu procjenu kompleksnosti algoritma, ako to nije pretesko. }
\subsubsection{Fitness function}\label{sec:fitness}
\fxnote{Marko: stigao do ovog dijela...}
Symbolic regression objective is to determine the expression governing available data. It is also allowed to obtain some equivalent expression, keeping in mind that there are multiple ways to express some symbolical equation. Although logically sound and intuitive, this objective is not quantifiable during the solution search/training phase, simply because the goal expression is not known at that point as only target values for some of the input features are known. Thus, different numerical metrics are being used in literature to guide the symbolic regression search process. The most popular are the coefficient of determination, also known as $R^2$ and mean squared error (MSE) or root mean squared error (RMSE). The important aspect of the solution quality might also be the solution expression complexity. This follows the Occam's razor principle~\cite{costa2020fast}~ that simpler solution  is more likely to be a correct one. 
The search process of \textsc{Rils}-\textsc{Rols}  is guided by the non-linear combination of $R^2$, RMSE and solution expression size (complexity) presented in Equation~\ref{eq:fitness}.

@@

Since the presented fitness function needs to be minimized, the following conclusions may be drawn:
\begin{itemize}
	\item higher $R^2$ is preferred -- ideally, when $R^2(s)=1$, the effect of term $1+R^2(s)$ is neutralized; %since 1 is neutral for multiplication;
	\item lower RMSE is preferred, ideally, when $RMSE(s)=0$, the whole term $(1+RMSE(s))$ becomes 1;
	\item since $penalty_{size} > 0$, larger expressions tend to have higher fitness (which follows the Occam's razor principle); therefore, we award solutions that are more simple with its representation. 
\end{itemize}

@@

Perturbations are generated by making simple changes on the per-node level of $s$ expression tree.
Depending on the structure of the expression tree (expression does not have to have unique tree representation), the set of subtrees of the previous expression $(x+y^3)\cdot x/\sqrt{y}$ might be $\{(x+y^3)\cdot x/\sqrt{y}, (x+y^3), x/\sqrt{y}, x, y^3, \sqrt{y}, y\}$. This set of subtrees is obtained if left subtree of the whole expression is $(x+y^3)$ while the right one is $x/\sqrt{y}$. 
For each subtree $n$, the set of perturbations around $n$ is generated (lines \ref{line:forSSStart}-\ref{line:forSSEnd} of Algorithm~\ref{alg:pert}).

Algorithm~\ref{alg:pertNode} shows how perturbations around given subtree (node) are generated.

@@

@@

	\begin{table}
		\centering
		\begin{tabular}{l}   \hline
			$x = \sqrt{x_1^2 + x_2^2 - 2 x_1 x_2 \cos(\theta_1 - \theta_2)}$ \\
			$ \theta_1 = \arcsin(n \sin \theta_2)$ \\
			$E =  \frac{m c^2 }{1 - \frac{v^2}{c^2}}$ \\
			$\omega = \frac{1 + \frac{v}{c}}{ \sqrt{1 - \frac{v^2}{c^2}}} \omega_0$ \\ \hline

>>>>

\fxnote{Dati neku smislenu procjenu kompleksnosti algoritma, ako to nije pretesko. }
\subsubsection{Fitness function}\label{sec:fitness}

Symbolic regression objective is to determine the expression fitting available data. It is also allowed to obtain some equivalent expression, keeping in mind that there are multiple ways to express some symbolical equation. Although logically sound and intuitive, this objective is not quantifiable during the solution search/training phase, simply because the goal expression is not known at that point as only target values for some of the input features are known. Thus, different numerical metrics are being used in literature to guide the symbolic regression search process. The most popular are the coefficient of determination, also known as $R^2$ and mean squared error (MSE) or root mean squared error (RMSE). The important aspect of the solution quality might also be the solution expression complexity, which may correspond to the model size of its tree representation. This follows the Occam's razor principle~\cite{costa2020fast}~ that simpler solution  is more likely to be a correct one. 
The search process of \textsc{Rils}-\textsc{Rols}  is guided by the non-linear combination of $R^2$, RMSE and solution expression size (complexity) presented in Equation~\ref{eq:fitness}.

@@

Since the presented fitness function needs to be minimized, the following conclusions may be drawn:
\begin{itemize}
	\item higher $R^2$ is preferred -- ideally, when $R^2(s)=1$, the effect of term $1+R^2(s)$ is neutralized; %since 1 is neutral for multiplication;
	\item lower RMSE is preferred, ideally, when $RMSE(s)=0$, the whole term $(1+RMSE(s))$ becomes 1;
	\item since $penalty_{size} > 0$, larger expressions tend to have higher fitness (which follows the Occam's razor principle); therefore, we award those solutions with a more simple representation. 
\end{itemize}

@@

Perturbations are generated by making simple changes on the per-node level of $s$ expression tree.
Depending on the structure of the expression tree (note that expression does not have to have unique tree representation), the set of subtrees of the previous expression $(x+y^3)\cdot x/\sqrt{y}$ might be $\{(x+y^3)\cdot x/\sqrt{y}, (x+y^3), x/\sqrt{y}, x, y^3, \sqrt{y}, y\}$. %This set of subtrees is obtained if left subtree of the whole expression is $(x+y^3)$ while the right one is $x/\sqrt{y}$.  
For each subtree $n$, the set of perturbations around $n$ is generated (lines \ref{line:forSSStart}-\ref{line:forSSEnd} in Algorithm~\ref{alg:pert}).

Algorithm~\ref{alg:pertNode} shows how perturbations around given subtree are generated.

@@

 \fxnote{Marko: stigao do ovog dijela...}

@@

	\begin{table}[!ht]
		\centering
		\begin{tabular}{l}   \hline
			$x = \sqrt{x_1^2 + x_2^2 - 2 x_1 x_2 \cos(\theta_1 - \theta_2)}$ \\
			$ \theta_1 = \arcsin(n \sin \theta_2)$ \\
			$E =  \frac{m c^2 }{1 - \frac{v^2}{c^2}}$ \\
			$\omega = \frac{1 + \frac{v}{c}}{ \sqrt{1 - \frac{v^2}{c^2}}} \omega_0$ \\ \hline
